title: "Medicare Fraud"
output: html_document
date: "2024-09-17"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(httr)
library(jsonlite)
library(sparklyr)
library(tidycensus)
library(dplyr)
library(purrr)
library(DBI)
library(knitr)
library(RMySQL)
library(data.table)
options(rstudio.connectionObserver.errorsSuppressed = TRUE)
```
The Data was too big to upload directly to R, so it was uploaded to mysql Workbench instead for aggregation before being pulled back into R

```{r}
con <- dbConnect(RMySQL::MySQL(), 

                 dbname = "fraud", 

                 host = "localhost", 

                 port = 3306, 

                 user = "root", 

                 password = "Kidace909")

dbSendQuery(con, "SET GLOBAL local_infile = true;") # <--- Added this

```




```{r}
# Open the file connection
cons <- file("C:\\Users\\bobbt\\Downloads\\Medicare_Part_D_Prescribers_by_Provider_and_Drug_2022.csv", open = "r")

# Count total lines in the file (each line corresponds to a row)
total_rows <- 0
while (length(line <- readLines(cons, n = 1, warn = FALSE)) > 0) {
  total_rows <- total_rows + 1
}
close(cons)

# Print total number of rows (including the header)
total_rows
```


```{r}
# Define the split point (half of the file)
split_point <- total_rows / 2

# Open connections for reading and writing
input_file <- file("C:\\Users\\bobbt\\Downloads\\Medicare_Part_D_Prescribers_by_Provider_and_Drug_2022.csv", open = "r")
output_file <- file("C:\\Users\\bobbt\\Downloads\\Medicare_Part_D_Second_Half.csv", open = "w")

# Read and write the header to the second half file
header <- readLines(input_file, n = 1)
writeLines(header, output_file)

# Skip the first half of the rows
for (i in 2:split_point) {
  readLines(input_file, n = 1)  # Skip rows without saving them
}

# Write the second half of the rows to the new file
for (i in (split_point + 1):total_rows) {
  line <- readLines(input_file, n = 1)
  writeLines(line, output_file)
}

# Close connections
close(input_file)
close(output_file)

```


read medicare_part_d and leie dataset and uploiad it to 
```{r}
# Define the file path
file_path <- "C:\\Users\\bobbt\\Downloads\\medicare_part_2.csv"
leie_path <- "C:\\Users\\bobbt\\Downloads\\leie.csv"

# Read the CSV file
data <- fread(file_path)
datas <- fread(leie_path)

# Check the number of rows
num_rows <- nrow(data)
print(num_rows)

dbWriteTable(con, "medicare_part_2_data", as.data.frame(data), overwrite = TRUE, row.names = FALSE)
dbDisconnect(con)
```



```{r}
dbWriteTable(con, "medicare_part_2_data", as.data.frame(data), overwrite = TRUE, row.names = FALSE)
dbDisconnect(con)
dbWriteTable(con, "leie", as.data.frame(datas), overwrite = TRUE, row.names = FALSE)
```


```{r}
dbWriteTable(con, "leie", as.data.frame(datas), overwrite = TRUE, row.names = FALSE)

```
Import the data into data frames
```{r}
# Import medicare_part_b
medicare_part_b <- dbGetQuery(con, "SELECT * FROM medicare_part_b")

# Import medicare_part_d
medicare_part_d <- dbGetQuery(con, "SELECT * FROM medicare_part_d")

# Import dmepos
dmepos <- dbGetQuery(con, "SELECT * FROM dmepos")

# Import combined_dataset
combined_dataset <- dbGetQuery(con, "SELECT * FROM combined_dataset")
```

Now lets call a function to calculate the percentage of fraudulent to non-fraudulent cases
```{r}
calc_fraud_stats <- function(dataset){
  total <- nrow(dataset)
  fraud <- sum(dataset$Exclusion == 1, na.rm = TRUE)
  non_fraud = total - fraud
  percent_fraud = (fraud/total) * 100
  return(c(Non_fraudulent = non_fraud,
           Fraudulent = fraud,
           Percent_Fraudulent= percent_fraud))
}

part_b_stats <- calc_fraud_stats(medicare_part_b)
part_d_stats <- calc_fraud_stats(medicare_part_d)
dmepos_stats <- calc_fraud_stats(dmepos)
combined_stats <- calc_fraud_stats(combined_dataset)

fraud_table <- data.frame(Dataset = c("Part B", "Part D", "DMEPOS", "Combined"),
  Non_fraudulent = c(part_b_stats["Non_fraudulent"], 
                     part_d_stats["Non_fraudulent"], 
                     dmepos_stats["Non_fraudulent"], 
                     combined_stats["Non_fraudulent"]),
  Fraudulent = c(part_b_stats["Fraudulent"], 
                 part_d_stats["Fraudulent"], 
                 dmepos_stats["Fraudulent"], 
                 combined_stats["Fraudulent"]),
  Percent_Fraudulent = c(part_b_stats["Percent_Fraudulent"], 
                         part_d_stats["Percent_Fraudulent"], 
                         dmepos_stats["Percent_Fraudulent"], 
                         combined_stats["Percent_Fraudulent"]))

fraud_table$Percent_Fraudulent <- round(fraud_table$Percent_Fraudulent, 3)
kable(fraud_table, format = "markdown", digits = 3)
```


